### 2021-04-14 newvariant.tex
rewriting the entire section

When the assumed distribution matches the true distribution, the estimated reproduction numbers match the true values (\fref{Rtbias}A), and therefore the ratio remains constant (\fref{Rtbias}B).
However, if the generation-interval distribution of the variant differs from the assumed distribution, the ratio changes over time (\fref{Rtbias}E,H).
In particular, if the true generation intervals have a shorter mean than the assumed distribution, we over-estimate $\Rv(t)$ during the growth phase and under-estimate during the decay phase (and vice versa, \fref{Rtbias}D,G), which further translates to biases in the estimated relative strength (\fref{Rtbias}F,I).

We can also assess these biases by plotting the estimated strength of the variant $\hat{\RR}_{\textrm{var}}(t)$ against the estimated strength of the wild type $\hat{\RR}_{\textrm{wt}}(t)$ (\fref{Rtbias}C,F,I).
When generation-interval distributions are correctly specified, we obtain a straight line with a slope of $\rho$ and intercept at zero (\fref{Rtbias}C).
When the assumed mean generation interval is longer than the that of the variant, we over-estimate the slope (and vice versa, \fref{Rtbias}F,I).
In addition, the intercept (if we were to perform regression) need not be at zero due to biases in the slope.
Smooth changes in $\RR$ further generate circular patterns but biases in slopes are clear.

@@

We consider two types of intervention:
an intervention of constant strength, which reduces transmission by a constant factor $\theta$ regardless of age of infection ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$); and an intervention of constant speed, which reduces transmission after infection by a constant rate $\phi$ ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$).
In this case, we can control the spread of the variant when $\theta > \Rv$ or $\phi > \rv$, respectively.
In general, we can calculate the strength ($\theta = \Ry{pre}/\Ry{post}$) and speed ($\phi = \ry{pre} - \ry{post}$) for any intervention by comparing epidemiological conditions before and after the intervention \citep{doi:10.1098/rspb.2020.1556}---these quantities, by definition, must be greater than the pre-intervention strength or speed of the epidemic in order to control the epidemic.
Here, we use the strength-and-speed framework to compare how differences in generation-interval distributions would affect the control of the spread of a new variant.

@@

\fref{strengthspeed} compares the effectiveness of strength-based and speed-based interventions across a wide range of assumptions about generation-interval distributions.
We begin by assuming that the variant and the wild type have equal mean generation intervals and model interventions such that they both reduce $\Rv$ to 0.9 after intervention (\fref{strengthspeed}A).
If the mean generation interval of the variant is longer than that of the wild type, then the strength of the variant will be higher;
since the strength of the strength-like intervention is insensitive to the underlying generation-interval distribution, we may fail to control the spread of the new variant.
With speed-like interventions, however, there is a countervailing effect:
since speed-like interventions like contact tracing have disproportionate effects on preventing later transmission, they are more effective (therefore stronger) when the mean generation interval of the new variant is longer.

The speed-based paradigm gives the same results about the control but provides important insights (\fref{strengthspeed}B).
The observed speed of the variant at any given moment is independent of our estimates of its mean generation interval.
Therefore, if we can reduce infectiousness at a faster rate than the observed speed of spread (i.e., if the speed of intervention is faster than the epidemic speed), we can control the epidemic---this explains why a constant-speed intervention can prevent the spread of the variant regardless of its mean generation interval.

>>>>

When the assumed distribution matches the true distribution, the estimated reproduction numbers match the true values (\fref{Rtbias}A), and therefore the ratio remains constant (\fref{Rtbias}B).
However, if the generation-interval distribution of the variant differs from the assumed distribution, the ratio changes over time (\fref{Rtbias}E,H).
In particular, if the true generation intervals have a shorter mean than the assumed distribution, we over-estimate $\Rv(t)$ during the growth phase and under-estimate during the decay phase (and vice versa, \fref{Rtbias}D,G), which further translates to biases in the estimated relative strength (\fref{Rtbias}E,H).
Therefore, biases in the estimated relative strength will correlate with the strength of the variant, (i.e., whether $\Rv(t)$ is below or above 1).

In practice, estimates of instantaneous reproduction numbers $\RR(t)$ (and therefore, their ratios) can be noisy due to limited data availability or model assumptions;
instead, we might want to estimate a single value of relative strength $\rho$.
For example, we can estimate $\rho$ by plotting the estimated strength of the variant $\hat{\RR}_{\textrm{var}}(t)$ against the estimated strength of the wild type $\hat{\RR}_{\textrm{wt}}(t)$---as presented in Figure 2 of \cite{volz2021transmission}---and performing a linear regression (\fref{Rtbias}C,F,I).
When generation-interval distributions are correctly specified, we obtain a straight line with a slope of $\rho$ and intercept at zero (\fref{Rtbias}C).
However, when the assumed mean generation interval is longer than the that of the variant, we over-estimate the slope (and vice versa, \fref{Rtbias}F,I).
In theory, we expect the regression line to go through the origin, but when the assumed generation-interval distribution differs from the true distribution, this is not necessarily the case: biases in slopes due to mis-specified generation-interval distributions also lead to biases in intercepts.

@@

We consider two types of intervention:
an intervention of constant strength, which reduces transmission by a constant factor $\theta$ regardless of age of infection ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$); and an intervention of constant speed, which reduces transmission after infection by a constant rate $\phi$ ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$).
In this case, we can control the spread of the variant when $\theta > \Rv$ or $\phi > \rv$, respectively.
As a baseline scenario, we consider constant-strength and speed interventions that reduce $\Rv$ to 0.9 when both the variant and the wild type have equal mean generation intervals (\fref{strengthspeed}A--B).
While both interventions are equally effective on strength-scale (that is, $\Ry{post}=\int  K_{\mathrm{post}}(\tau) \dtau = 0.9$), they have different dynamical implications.
The constant-strength intervention affects transmission equally throughout the course of infection, whereas the constant-speed intervention has greater impact of transmission that occur later;
as a result, the constant-speed intervention reduces the post-intervention mean generation interval (\fref{strengthspeed}B) and leads to (slightly) faster exponential decay (therefore, lower $\ry{post}$).

@@

However, if the variant has longer generation intervals than the wild type (\fref{strengthspeed}C--D),  then the strength of the variant will be higher conditional on the observed speed (\fref{strengthspeed}E).
In this case, the same constant-strength intervention can fail to control the epidemic (i.e., $\Ry{post} > 1$; \fref{strengthspeed}E) because this intervention reduces the transmission by a constant amount regardless of age of infection (\fref{strengthspeed}C).
On the other hand, the same constant-speed intervention will prevent a larger proportion of transmission, leading to lower $\Ry{post}$ (\fref{strengthspeed}E), because it is more effective against late-stage transmission (\fref{strengthspeed}D).
The constant-speed intervention also has a greater impact on reducing the mean generation interval (\fref{strengthspeed}D).

The speed-based paradigm gives the same results about the control but provides important insights (\fref{strengthspeed}F).
The observed speed of the variant $\rv$ at a given moment is independent of our estimates of its mean generation interval.
Likewise, the post-intervention speed of the variant under the constant-speed intevention is also independent of the mean generation interval;
therefore, the constant-speed intevention will be always equally effective (on the speed scale) in preventing the spread.

The constant post-intervention speed can be understood by considering a speed-based decomposition of the kernel: $K_{\mathrm{pre}}(\tau) = \exp(\rv) b(\tau)$, where $b(\tau)$ is the initial backward generation-interval distribution and therefore integrates to 1 \citep{doi:10.1098/rspb.2020.1556}.
This decomposition allows us to re-write the post-intervention kernel (under the constant-speed intervention) as follows:
\begin{equation}
K_{\mathrm{post}}(\tau) = \exp(\rv-\phi) b(\tau).
\end{equation}
Since the post-intervention speed $\ry{post}$ is determined by the following implicit equation,
\begin{equation}
1 = \int_0^\infty \exp(-\ry{post} \tau) K_{\mathrm{post}}(\tau) \dtau = \int_0^\infty \exp(\rv-\phi-\ry{post}) b(\tau)  \dtau,
\end{equation}
we have $\ry{post} = \phi - \rv$. Therefore, if we can reduce infectiousness at a faster rate than the observed speed of spread (i.e., if the speed of intervention is faster than the epidemic speed, $\phi > \rv$), we can control the epidemic regardless of the underlying generation-interval distribution.
